Adding selection to \stdpopsim: oh how the tables have turned

### 2024-05-29 analysis2_manuscript.tex
working on results and fig 1

    % stdpopsim
    A lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers. This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community.
    While that is so, the current version of \stdpopsim does not include
    models of selection, which is a major limitation for many applications
    of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding the processes such
    as adaptation, the effects of selective sweeps, and the impact of
    background selection on genetic diversity. Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    accessible to a wide range of researchers and highly reproducible. 
    Further the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the density of functional elements, and relevant
    population size histories.

>>>>

    % stdpopsim
    A lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[][e.g.,]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
    While that is so, the current version of \stdpopsim does not include
    models of selection, which is a major limitation for more empirical
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding the processes such
    as adaptation, the effects of selective sweeps, and the impact of
    background selection on genetic diversity. Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Further the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the density of functional elements, and relevant
    population size histories.

    % goals
    In this manuscript we provide an overview of a major new addition
    to \stdpopsim-- the inclusion of models of selection.
    We begin by describing the models of selection that we have implemented,
    as well as the parameter choices that are available to the user.
    This includes models of background selection, selective sweeps, and
    models of the distribution of fitness effects.
    Further we describe how these models can be combined with genomic
    annotations that are specific to a given species available
    through the \stdpopsim API, and combined with a parameterized model of
    demography and recombination to provide a complete simulation of
    genetic variation in a population.
    We then provide a series of examples of how these models can be used
    to benchmark and compare the performance of different methods for
    inferring demographic history, selection, and the distribution of
    fitness effects from genetic data.

\section*{Implementing Selection in \stdpopsim}
    \label{selection}

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allowed for curated addition of new
    species and models, and relied on backend simulation engines,
    including msprime \citep{Baumdicker2022} and SLiM \citep{haller2019slim},
    for efficient simulation of genetic variation.
    To extend this to include models of selection, 
    we introduce an important new class representing a specfic distribution of fitness effects
    to the \stdpopsim API, the \texttt{DFE} class. 
    This class provides an interface for specifying
    a distribution of the fitness effects of new mutations, and
    is added to a \texttt{Contig} object to provide a description
    of the fitness effects of mutations along a chromosome.
    A \texttt{DFE} object can apply to a specific interval set of a chromosome,
    and can be combined with other \texttt{DFE} objects to provide, 
    a rich model of how selection may differ along a chromosome. 
    Once specified a model of selection is then run in the same way
    but currently is limited to using SLiM simulation engine, which
    is probably the most flexible simulation engine for modeling selection currently available.

    As with other \stdpopsim models, we aim to hew as closely as possible
    to realistic genomes, and provide public annotations of specices-specific functional
    elements that can be used along with a \texttt{DFE} object to
    readily provide realistic model of selection for a specific genome and set of
    populations. 
    % A figure here would be great to show how this works

    % current numbers of DFEs / species / etc
    the \stdpopsim
